Comparison among doublet predictions on benchmarking datasets

### 2024-06-25 analyses/doublet-detection/exploratory-notebooks/03_compare-benchmark-results.Rmd
clean up PCA code

plot_pca_metrics <- function(df, color_column, metric_colors) {
  # Plot PCs colored by performance metric, showing false calls on top
  # metric_colors is a named vector of colors used for coloring tp/tn/fp/fn
  # df is expected to contain columns PC1, PC2, and `color_column`. This should _not_ be provided as a string.
  ggplot(df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(
    size = 0.75, 
    alpha = 0.6
  ) + 
  geom_point(
    data = dplyr::filter(df, {{color_column}} %in% c("fp", "fn")), 
    size = 0.75
  ) +
  scale_color_manual(name = "Call type", values = metric_colors) +
  ggtitle("Consensus call metrics")
}
```

@@

      # Add consensus columns:
      # - consensus_class: "doublet" if all three methods are doublet, "singlet" if all three methods are singlet, 
      #    and "ambiguous" if there are any disagreements
      # - consensus_call: "doublet" if all three methods are doublet, and "singlet" otherwise
      # - confusion_call: tp/tn/fp/fn based on the `consensus_call` column
      dataset_df <- dataset_df |>
        dplyr::rowwise() |>
        # Add column `consensus_class`
        dplyr::mutate(
          consensus_class = dplyr::case_when(
            consensus_call == "doublet" ~ "doublet",
            all(c(scdbl_prediction, scrublet_prediction, cxds_prediction) == "singlet") ~ "singlet", 
            .default = "ambiguous"
          )
         ) |>
        # Add column `consensus_call`
        dplyr::mutate(
          consensus_call = dplyr::if_else(
            all(c(scdbl_prediction, scrublet_prediction, cxds_prediction) == "doublet"),
            "doublet", 
            "singlet"
          )
        ) |>
        # Add column `confusion_call`
        dplyr::mutate(
          confusion_call = dplyr::case_when(
            consensus_call == "doublet" && ground_truth == "doublet" ~ "tp",
            consensus_call == "singlet" && ground_truth == "singlet" ~ "tn",
            consensus_call == "doublet" && ground_truth == "singlet" ~ "fp",
            consensus_call == "singlet" && ground_truth == "doublet" ~ "fn"
          )
        )

@@

      # Top row: methods themselves ---
      p1 <- plot_pca_calls(
        df, 
        color_column = scdbl_prediction, 
        title = "scDblFinder prediction"
      )

      p2 <- plot_pca_calls(
        df, 
        color_column = scrublet_prediction, 
        title = "scrublet prediction"
      )

      p3 <- plot_pca_calls(
        df, 
        color_column = cxds_prediction, 
        title = "cxds prediction"
      )

      # Bottom row: consensus  --
      p4 <- plot_pca_calls(
        df, 
        color_column = ground_truth, 
        title = "Ground truth"
      )

      # Second, consensus call
      p5 <- plot_pca_calls(
        df, 
        color_column = consensus_call, 
        title = "Consensus call"
      )

      # Third, call type
      p6 <- plot_pca_metrics(
        df,
        call_type, 
        metric_colors = confusion_colors
      )

      # combine and plot
      plot( 
        (p1 + p2 + p3) / (p4 + p5 + p6) + 
          plot_annotation(
            glue::glue("PCA for {dataset}"), 
            theme = theme(plot.title = element_text(size = 16))) + 
          plot_layout(guides = "collect"))
    }
  )
```

@@

This section calculates a confusion matrix and associated statistics on the consensus calls.

@@

        cat("Table of consensus calls:")
        print(table(df$consensus_call))

>>>>

plot_pca_metrics <- function(df, color_column, metric_colors) {
  # Plot PCs colored by performance metric, showing false calls on top
  # metric_colors is a named vector of colors used for coloring tp/tn/fp/fn
  # df is expected to contain columns PC1, PC2, and `color_column`. This should _not_ be provided as a string.
  ggplot(df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(
    size = 0.75, 
    alpha = 0.6
  ) + 
  geom_point(
    data = dplyr::filter(df, {{color_column}} %in% c("fp", "fn")), 
    size = 0.75
  ) +
  scale_color_manual(name = "Call type", values = metric_colors) +
  ggtitle("Consensus class metrics")
}
```

@@

      # Add consensus columns:
      # - consensus_class: "doublet" if all three methods are doublet, "singlet" if all three methods are singlet, 
      #    and "ambiguous" if there are any disagreements
      # - consensus_call: "doublet" if all three methods are doublet, and "singlet" otherwise
      # - confusion_call: tp/tn/fp/fn based on the `consensus_call` column
      dataset_df <- dataset_df |>
        dplyr::rowwise() |>
        # Add column `consensus_call`
        dplyr::mutate(
          consensus_call = dplyr::if_else(
            all(c(scdbl_prediction, scrublet_prediction, cxds_prediction) == "doublet"),
            "doublet", 
            "singlet"
          )
        ) |>
        # Add column `consensus_class`
        dplyr::mutate(
          consensus_class = dplyr::case_when(
            consensus_call == "doublet" ~ "doublet",
            all(c(scdbl_prediction, scrublet_prediction, cxds_prediction) == "singlet") ~ "singlet", 
            .default = "ambiguous"
          )
         ) |>
        # Add column `confusion_call`
        dplyr::mutate(
          confusion_call = dplyr::case_when(
            consensus_call == "doublet" && ground_truth == "doublet" ~ "tp",
            consensus_call == "singlet" && ground_truth == "singlet" ~ "tn",
            consensus_call == "doublet" && ground_truth == "singlet" ~ "fp",
            consensus_call == "singlet" && ground_truth == "doublet" ~ "fn"
          )
        )

@@

      pc_color_columns <- c("scDblFInder prediction" = "scdbl_prediction",
                            "scublet prediction" = "scrublet_prediction",
                            "cxds prediction" = "cxds_prediction",
                            "Ground truth" = "ground_truth", 
                            "Consensus call" = "consensus_call")

      pc_plot_list <- pc_color_columns |> 
        purrr::imap(
          \(color_column, plot_title) {
            plot_pca_calls(
              df, 
              color_column = !!sym(color_column), 
              title = plot_title
            )}
        )

      # note that the list name isn't used here
      pc_plot_list$metrics <- plot_pca_metrics(
        df,
        confusion_call, 
        metric_colors = confusion_colors
      )

      patchwork::wrap_plots(pc_plot_list, pca_metrics) +
        plot_annotation(
            glue::glue("PCA for {dataset}"), 
            theme = theme(plot.title = element_text(size = 16))) + 
          plot_layout(guides = "collect") 
    }
  )
```

@@

This section calculates a confusion matrix and associated statistics on the consensus classes.

@@

        cat("Table of consensus calls:")
        print(table(df$consensus_class))
